---
title: "Practical Machine Learning - Course Project"
output: html_document
---

```{r}
library(caret)
```

This report is about how to predict the kind of an excersize by analyzing the movement data. The report fits a linear model and a random forrest and compares the performance of both models. The better model is taken to predict the values for submission.

# Load and Prepare the Data

## Load the Data
Download the dataset and load it in memory.
```{r}
pml_training_data <- read.csv("data/pml-training.csv", na.strings=c("NA","#DIV/0!",""), stringsAsFactors = TRUE)
```
## Clean the Data
The first seven rows are removed as they are id columns and data to identify groups of rows.

```{r}
# Remove id columns
trainingData <- pml_training_data[, -c(1:7)]

# remove columns with a variance near zero
nearZeroVarColumns <- nearZeroVar( trainingData )
trainingData <- trainingData[, -nearZeroVarColumns]

# Remove columns which contains more than 97 percent NAs
naColumns <- sapply(trainingData, function(x) mean(is.na(x))) > 0.97
trainingData <- trainingData[, naColumns == FALSE]
```


## Split the data in a training set and a testing set.

```{r}
inTrain <- createDataPartition(y=trainingData$classe, p=0.6, list=FALSE)

training <- trainingData[inTrain,]
testing  <- trainingData[-inTrain,]
```

# Preprocessing

The fitting of the linear model works better if the values are centered (substracted by the mean) and scaled (divided by the standard deviation).
```{r}
preProcValues <- preProcess(training, method = c("center", "scale"))
scaledTrain   <- predict(preProcValues, training)
scaledTesting <- predict(preProcValues, testing)
```

# Fit Model

The strategy to fit two models, a simple linear model and a random forrest. For both models the out of sample error rate is calculated using cross validation. Finally the models are validated with the testing data.

There are two models fitted because linear models work quite well in many cases and are simple to interpret. Random forrest models perform very good if there are many variables and the data contain none linear relationships.

The models are validated with cross validation. This method provides a good estimate about how the model will perform on new data and shows overfitting.

## CART Model
```{r}
set.seed(9764)
fitControl <- trainControl(method = "cv", number = 3, savePredictions = TRUE)

rpartFit <- train(classe ~ ., data = scaledTrain, method = "rpart", trControl = fitControl)
print(rpartFit)


predictions <- predict(rpartFit, newdata = scaledTesting)
cfm <- confusionMatrix(predictions, scaledTesting$classe)
print(cfm)

# Out of sample error rate
ooer <- 1 - as.numeric(cfm$overall[1])
ooer
```
The estimation of the out of sample error for this model is `r ooer` which is quite high. This may have to do with missing terms in the the formula.

## Random Forrest
```{r}
set.seed(9764)
fitControl <- trainControl(method = "cv", number = 3)

rfFit <- train(classe ~ ., data = scaledTrain, method = "rf", trControl = fitControl)
print(rfFit)


predictions <- predict(rfFit, newdata = scaledTesting)
cfm <- confusionMatrix(predictions, scaledTesting$classe)
print(cfm)

# Out of sample error rate
ooer <- 1 - as.numeric(cfm$overall[1])
ooer
```
The estimation of the out of sample error for this model is `r ooer`. This is a good result. Compared with a linear model, random forrest need no tuning of the formula and perform quite good out of the box.

# Conclussion

The random forrest model performs very good on the provided data. For the model only 53 columns of the data are used because the others contains not enough signal or unusable data.

# Create Submission

## Perform Data Cleaning
The data must be cleaned in the same way than the training set was cleaned.

```{r}
pml_testing_data  <- read.csv("data/pml-testing.csv", na.strings=c("NA","#DIV/0!",""), stringsAsFactors = TRUE)
# Remove id columns
testingData <- pml_testing_data[, -c(1:7)]
testingData <- testingData[, -nearZeroVarColumns]
testingData <- testingData[, naColumns == FALSE]

scaledTestSubmission   <- predict(preProcValues, testingData)
```
## Predict the submission results 
```{r}
submissionPredictions <- predict(rfFit, newdata = scaledTestSubmission)
```

## Write submission files
```{r}
 
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("submission/problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(submissionPredictions)
```